Leaf traits drive differences in biomass partitioning among major plant functional types

### 2015-07-21 manuscript.Rnw
References fixed; more added and text fixed in intro

\usepackage{amssymb,amsmath}
\usepackage{palatino}
\usepackage{framed}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage{fixltx2e}
\usepackage[default,osfigures,scale=0.95]{opensans}
\graphicspath{ {./figures/} }
\usepackage{geometry}
\usepackage{authblk}
\usepackage{gensymb}
\usepackage{longtable}
\usepackage{lineno}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}

@@

% paragraph on what we do know about differences between PFTs; what we might expect (LMF vs. LAR; LMA and leaf lifespan differences)
Previous work based on whole stands\citep{oneilldeangelis1981,enquistniklas2002,reich2014} and individual plants\citep{poorter2012}, reported differences between angiosperms and gymnosperms in the amount of leaf biomass per unit above-ground biomass (the 'leaf mass fraction', LMF). However, it is unclear whether PFTs with higher leaf biomass also have higher total leaf area. Leaf area is relevant because it drives total light interception and thus photosynthesis. Several classic studies have assumed, based only on differences in measurements of leaf mass, that gymnosperms carry more leaf area than angiosperms \citep{chabothicks1982, bond1989}.

%paragraph on LMF, it is an interesting variable because it is an important component of turnover of biomass. Link to leaf lifespan across PFTs as well. 
For the above-ground biomass pools, we calculated the leaf mass fraction ($M_F / M_T$, where $M_T$ is total above-ground biomass) and leaf area ratio ($A_F / M_T$). These variables are related by,

$$\textrm{LMF} = \frac{M_F}{M_T} = \left(\frac{M_F}{A_F}\right) \left(\frac{A_F}{M_T}\right) = \textrm{LMA} \times \textrm{LAR}.$$

For LMF we hypothesize that it is higher for EG than DA, with EA taking an intermediate value as determined by its mean LMA. This would be consistent both with leaf lifespan correlate as well as minimizing variation in leaf area per unit biomass (which determines light interception).

We also wonder whether differences in LMF and LAR (and its component variables) are sensitive to climate. One may expect this to be the case because short-term experiments have shown responsiveness of biomass partitioning to soil water etc. etc.; or this variation may wash out at a large scale. It is thus the question whether variation within PFT can be explained by climate or not. Indeed, many GVMs assume that climate affects biomass partitioning within a given PFT, often through soil water stress or other abiotic stress factors (see Methods \ref{sec:gvmsummary}).

Here, we use a unique, new database\citep{Falster2015} (Fig. 1) to establish general rules on how biomass partitioning differs among three dominant woody PFTs across the globe: evergreen gymnosperms, evergreen angiosperms, and deciduous angiosperms. The dataset includes over 21,000 observations of biomass and size metrics for individual plants, contributed by 175 studies across nine vegetation types (Fig. S1).

One key challenge is to account for the very large size variation in any allometric variable, also LMF. We do this by fitting an assumption-free model (GAM) and studying PFT patterns at a common height.

@@

We only used LMA directly estimated for the harvested plants (typically for a subsample of leaves). For conifers, leaf area was converted to half-total surface area using the average of a set of published conversion factors \citep{barclay2000}. We used a different conversion for pines (\emph{Pinus} spp.) than non-pines, because the conversion factors were very different for these two groups \citep{barclay2000}. This conversion was necessary because half-total surface area is most appropriate for comparison to flat leaves \citep{lang1991, chen1992}. Stem cross-sectional area was measured either at breast height (ca. 1.3m) and/or at the base of the plant. For the subset of the data where both were measured, we estimated basal stem area ($A_S$) from breast height stem area ($A_{Sbh}$) using a non-linear regression model (see~\ref{sec:predas}).

@@

For Eq.\ref{eq:M1} we need the basal diameter, but this is not normally measured for large plants. Using the subset of data where basal stem diameter ($A_S$) and diameter at breast height ($A_{Sbh}$) were measured (a total of \Sexpr{length(residuals(basalafit))} observations covering the three major PFTs), we developed a non-linear model to predict $A_S$ when only $A_{Sbh}$ and plant height ($H$) were measured. We fit the following equation, using non-linear regression.

@@

paragraph - Why would LAR be less variable than LMF? Because light interception. Lulu effect paper. Need to discuss linkage between LMA, leaf lifespan, leaf area vs. leaf mass at given MT (and given height!).

paragraph - variation in LMF was the same as MF/AS, suggesting that MT/AS is a fairly conserved quantity \ref{eq:M2}. Consistent with \citep{chave2005} who showed that aboveground biomass can be well estimated with Ht*AS*wood density. This works because, for large plants, most of aboveground biomass is wood.

>>>>

\usepackage{amssymb,amsmath}
\usepackage{palatino}
\usepackage{framed}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage{fixltx2e}
% \usepackage[default,osfigures,scale=0.95]{opensans}
\graphicspath{ {./figures/} }
\usepackage{geometry}
\usepackage{authblk}
\usepackage{gensymb}
\usepackage{longtable}
\usepackage{lineno}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}

@@

In this paper we are primarily interested in the distribution of biomass ('partitioning') between leaves and stems, an important component of the residence time of carbon in ecosystems \citep{friend2014}. Previous work based on whole stands \citep{oneilldeangelis1981,enquistniklas2002,reich2014} and individual plants \citep{poorter2012}, reported differences between angiosperms and gymnosperms in the amount of leaf biomass per unit above-ground biomass (the 'leaf mass fraction', LMF). \citet{poorter2012} and \citet{enquist2002} attributed higher LMF in gymnosperms to longer leaf lifespan (LL) compared to typical angiosperms, but this begs the question whether differences in LMF are also apparent between deciduous and evergreen functional types within angiosperms. It is also unknown whether PFTs with higher leaf biomass also have higher total leaf area, which is relevant because it drives total light interception and thus productivity. Several classic studies have assumed, based only on differences in measurements of leaf mass, that gymnosperms carry more leaf area than angiosperms \citep{chabothicks1982, bond1989}.

Little is known about global-scale patterns in LMF and LAR in relation to climate. One may expect that biomass partitioning is correlated with precipitation or mean annual temperature \citep{reich2014} because short-term experiments (cite) and climate gradients \citep{callaway1994,poyatos2007} have shown responsiveness of biomass partitioning and related traits. It is thus the question whether variation within PFT can be explained by climate or not. Indeed, many GVMs assume that climate affects biomass partitioning within a given PFT, often through soil water stress or other abiotic stress factors (see Methods \ref{sec:gvmsummary}).

Here, we use a unique, new database \citep{Falster2015} (Fig. 1) to establish general rules on how biomass partitioning differs among three dominant woody PFTs across the globe: evergreen gymnosperms, evergreen angiosperms, and deciduous angiosperms. The dataset includes observations of biomass and size metrics for individual plants, contributed by 175 studies across nine vegetation types (Fig. 1). In this paper we include field-grown plants only, including the entire size range of woody plants (0.01 - \textgreater 100m height). One key challenge is thus to account for the very large size variation commonly found in any allometric variable \citep{niklas1994}. We do this by fitting a statistical model that does no assume a functional dependence on plant size, allowing us to study PFT patterns at a common plant height.

@@

We only used LMA directly estimated for the harvested plants (typically for a subsample of leaves). For conifers, leaf area was converted to half-total surface area using the average of a set of published conversion factors \citep{barclay2000}. We used a different conversion for pines (\emph{Pinus} spp.) than non-pines, because the conversion factors were very different for these two groups \citep{barclay2000}. This conversion was necessary because half-total surface area is most appropriate for comparison to flat leaves \citep{lang1991, chen1992}. Stem cross-sectional area was measured either at breast height (ca. 1.3m) and/or at the base of the plant. For the subset of the data where both were measured, we estimated basal stem area ($A_S$) from breast height stem area ($A_{Sbh}$) using a non-linear regression model, as follows.

Using the subset of data where basal stem diameter ($A_S$) and diameter at breast height ($A_{Sbh}$) were measured (a total of \Sexpr{length(residuals(basalafit))} observations covering the three major PFTs), we developed a non-linear model to predict $A_S$ when only $A_{Sbh}$ and plant height ($H$) were measured. We fit the following equation, using non-linear regression.

@@

@@

Poorter et al 2012 did not find a clear relationship between LMF and LMA across 8 PFTs, perhaps not enough data? Also, we only include field-grown plants, they have everything. It also seems they did not account for plant size in that analysis.

paragraph - Why would LAR be less variable than LMF? Because light interception. Lulu effect paper. Need to discuss linkage between LMA, leaf lifespan, leaf area vs. leaf mass at given MT (and given height!).
One option is that LAR is targeted to be more constant, because it drives light interception. Very high leaf areas lead to very shaded foliage (etc.). Other option is that this arises through the correlation with leaf lifepan. This is how Enquist \& Niklas explained it, but it is only true when all else is equal. In support, \citet{reich1992} showed positive correlation between leaf life span and total stand leaf biomass across diverse forest stands. Our results are certainly consistent with longer leaf lifespan leading to higher leaf biomass, because of the strong nearly 1/x correlation between LL and LMA ()

paragraph - variation in LMF was the same as MF/AS, suggesting that MT/AS is a fairly conserved quantity \ref{eq:M2}. Consistent with \citep{chave2005} who showed that aboveground biomass can be well estimated with c*Ht*AS*wood density (works because geometry). This works because, for large plants, most of aboveground biomass is wood.
